---
title: "2.Data_intergration"
output: html_document
date: "2024-04-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r}
#devtools::install_github("immunogenomics/harmony") 
#install.packages("harmony")
```

```{r}
library(tidyverse)
library(Seurat)
library(patchwork)
library(SeuratData)
library(patchwork)
library(sctransform)
library(harmony)
set.seed(9999)
```

```{r}

# Load the datasets after 1st Qc and doublet finder! Columns are not removed yet !

ctl_1_df <- readRDS("SCT/SCT_transf/ctl_1_df_1stfilt_SCT.rds")
ctl_2_df <- readRDS("SCT/SCT_transf/ctl_2_df_1stfilt_SCT.rds")

int_1_df <- readRDS("SCT/SCT_transf/int_1_df_1stfilt_SCT.rds")
int_2_df <- readRDS("SCT/SCT_transf/int_2_df_1stfilt_SCT.rds") 
int_3_df <- readRDS("SCT/SCT_transf/int_3_df_1stfilt_SCT.rds")

end_1_df <- readRDS("SCT/SCT_transf/end_1_df_1stfilt_SCT.rds")
end_2_df <- readRDS("SCT/SCT_transf/end_2_df_1stfilt_SCT.rds")
end_3_df <- readRDS("SCT/SCT_transf/end_3_df_1stfilt_SCT.rds")

stages.names <- list("ctl_1_df", "ctl_2_df", "int_1_df","int_2_df", "int_3_df","end_1_df", "end_2_df", "end_3_df")

```

```{r}
for (name in stages.names) {
  
  current_object <- get(name)
  doublet_col <- colnames(current_object@meta.data)[13]
  doublet_table <- table(current_object@meta.data[doublet_col])
  umap_doublets <- DimPlot(current_object,reduction = "umap", group.by = doublet_col)
  print(name)
  print(doublet_table)
  print(umap_doublets)
  
  cat(name, "\n\n- Cells:", nrow(current_object@meta.data), "\n- Features:", nrow(current_object), "\n- Meta Cols:", ncol(current_object@meta.data), "\n")
}
```


## Remove the doublets based on the last column- important to run only once otherwise you keep removing columns! Also checks if the doublet column is present (based on ;doubletFinder!)

```{r}
for (name in stages.names) {
  current_object <- get(name)
  #remove pAnn and 1dt non adj classification
  current_object@meta.data <- current_object@meta.data[, -c(11, 12), drop = FALSE]
  
  # find the "DF.classifications_" column with the singlets,doublets
  doublet_col <- grep("^DF.classifications_", colnames(current_object@meta.data), value = TRUE)
  
  if (length(doublet_col) == 1) {
    target_col <- doublet_col[1]
    cells_to_keep <- rownames(current_object@meta.data)[current_object@meta.data[[target_col]] == "Singlet"]
  
    current_object <- subset(current_object, cells = cells_to_keep) # we subset the whole seurat based on the singlets
    
  } else if (length(doublet_col) == 0) {
    cat("No 'DF.classifications_' column found in", name, "\n")
  } else {
    cat("Multiple 'DF.classifications_' columns found in", name, "please discard and use one\n")
  }
  
  cat(name, "\n- Cells:", nrow(current_object@meta.data), "\n- Features:", nrow(current_object), "\n- Meta Cols:", ncol(current_object@meta.data), "\n")
  assign(name, current_object)
}
```

# work with list to make a merged object later on

```{r}
#RasV_stages.list <- list(ctl_1_df, ctl_2_df, int_1_df, int_2_df, int_3_df, end_1_df, end_2_df, end_3_df)
# try the later one
RasV_stages.list <- list(
  ctl_1_df = ctl_1_df,
  ctl_2_df = ctl_2_df,
  int_1_df = int_1_df,
  int_2_df = int_2_df,
  int_3_df = int_3_df,
  end_1_df = end_1_df,
  end_2_df = end_2_df,
  end_3_df = end_3_df
)

rm(ctl_1_df, ctl_2_df, int_1_df, int_2_df, int_3_df, end_1_df, end_2_df, end_3_df)
gc()
```

## SCT function on the new filtered seurat dataframes ## not sure if i should use the whole cell dataset

```{r}
# process pipeline
process_object <- function(seurat_object) {
  seurat_object %>%
    SCTransform(vars.to.regress = "percent.mt", ncells = ncol(seurat_object), verbose = TRUE) %>%
    RunPCA(assay = "SCT") %>%
    RunUMAP(dims = 1:50, assay = "SCT") %>%
    FindNeighbors(reduction = "pca", dims = 1:50) %>%
    FindClusters(algorithm = 4, method = "igraph", resolution = 0.07) %>%
    RunUMAP(dims = 1:50, assay = "SCT") 
}
```

## Apply the sct function in each dataset

```{r}
RasV_stages.list <- lapply(RasV_stages.list, process_object)

```

```{r}
pdf("UMAP_SCT_post_dblRem.pdf", width = 10, height = 10)


for (i in seq_along(RasV_stages.list)) {
  current_object <- RasV_stages.list[[i]]
  umap <- DimPlot(current_object, reduction = "umap", label = TRUE) 
  # Use names(RasV_stages.list)[i] to fetch the name of the current object
  umap_title <- umap + ggtitle(paste("UMAP_SCT_post_dblRem", names(RasV_stages.list)[i]))
  print(umap_title)
}

dev.off()

```

```{r}
for (i in seq_along(RasV_stages.list)) {
  current_object <- RasV_stages.list[[i]]
  name <- names(RasV_stages.list)[i]
  filepath <- paste0("output/", name, "sct_dblt_filt.rds")
  saveRDS(current_object, file = filepath)
}
```


########################### MERGING PART ####################################

```{r}
# import again the sct doublet cleaned datasets if needed
ctl_1_df <- readRDS("SCT/doublet_removed/ctl_1_dfsct_dblt_filt.rds")
ctl_2_df <- readRDS("SCT/doublet_removed/ctl_2_dfsct_dblt_filt.rds")

int_1_df <- readRDS("SCT/doublet_removed/int_1_dfsct_dblt_filt.rds")
int_2_df <- readRDS("SCT/doublet_removed/int_2_dfsct_dblt_filt.rds")
int_3_df <- readRDS("SCT/doublet_removed/int_3_dfsct_dblt_filt.rds")

end_1_df <- readRDS("SCT/doublet_removed/end_1_dfsct_dblt_filt.rds")
end_2_df <- readRDS("SCT/doublet_removed/end_2_dfsct_dblt_filt.rds")
end_3_df <- readRDS("SCT/doublet_removed/end_3_dfsct_dblt_filt.rds")

RasV_stages.list <- list(
  ctl_1_df = ctl_1_df,
  ctl_2_df = ctl_2_df,
  int_1_df = int_1_df,
  int_2_df = int_2_df,
  int_3_df = int_3_df,
  end_1_df = end_1_df,
  end_2_df = end_2_df,
  end_3_df = end_3_df
)



```

```{r}
rm(ctl_1_df, ctl_2_df, int_1_df, int_2_df, int_3_df, end_1_df, end_2_df, end_3_df)
gc()
```


```{r}
# Add new column as a single identifier for each dataset that we have by cobining replicate number and condition named batch
for (i in seq_along(RasV_stages.list)) {

  if ("replicate" %in% colnames(RasV_stages.list[[i]]@meta.data)) {
    replicate_info <- RasV_stages.list[[i]]@meta.data$replicate
  } else {
    replicate_info <- "NA"
  }
  
  # we assume condition column always exists (if we use my Qc pipeline)
  condition_info <- RasV_stages.list[[i]]@meta.data$condition

  # new id column 'batch'
  RasV_stages.list[[i]]@meta.data$batch <- paste(condition_info, replicate_info, sep = "_")

}
```

########### Use this only if you dont SCT the global object!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
```{r}

integ_features <- SelectIntegrationFeatures(RasV_stages.list, nfeatures = 3000)
```

## Merge the objects into 1


# rename cell ids

```{r}
# rename cells, base d on the batch identifier, so we dont have doublicates when we merge
for (i in seq_along(RasV_stages.list)){
  RasV_stages.list[[i]] <- RenameCells(RasV_stages.list[[i]], add.cell.id = unique(RasV_stages.list[[i]]@meta.data$batch))
}
```

# merge seurat objects

```{r}
## some names seems to be doublicated across the objects so in previous part i remake the id
RasV_merged <- merge(
  x = RasV_stages.list[[1]],               # The first object
  y = RasV_stages.list[-1],            # A vector of additional Seurat objects to merge , merged object seems to be a bit larger but maybe merge add more information
  project = "NASH_HCC",
  merge.data = TRUE
)
DefaultAssay(RasV_merged) <- "SCT" #important if we want to change the default option of the preprocessing steps
```

# A ) workflow of process seurat global object if needed, i probably dont need THIS!!!!!!!!

```{r}
# normalize and PCA
#VariableFeatures(RasV_merged[["SCT"]]) <- rownames(RasV_merged[["SCT"]]@scale.data)
RasV_merged <-  SCTransform(RasV_merged, vars.to.regress = "percent.mt", verbose = FALSE)##!!!!!!!!!!!! change the number of cells to 3000-default in the normal run!!!!!!!!!!!!!!
RasV_merged <-  RunPCA(RasV_merged, assay = "SCT")
RasV_merged <-  RunUMAP(RasV_merged, dims = 1:50, assay = "SCT")

```
### B) Continue with the non global transformation just with individual SCT merged objects and look for batch effects

```{r}
VariableFeatures(RasV_merged) <- integ_features

# Calculate PCs using manually set variable features
RasV_merged <- RunPCA(RasV_merged, assay = "SCT", npcs = 50)
RasV_merged <- RunUMAP(RasV_merged,dims = 1:50, assay = "SCT" )
DimPlot(RasV_merged,reduction = "umap", group.by = "batch", shuffle = TRUE)
```


```{r}
# save the merged file in case the programm crashes again
saveRDS(RasV_merged, file = "SCT/merged_individualy_processed.rds")
```


## HERE we run harmony embedding (batch correction of pca)


```{r}
RasV_merged <- readRDS("SCT/processed_merged.rds")

RasV_merged <- RunHarmony(RasV_merged, group.by.vars = "batch")  # harmony uses the PCA embeddings of the previous normalization , use :c("batch", "condition"), just dataset id-batch performs betetr on clustering together replicates and make conditions more distinct
RasV_merged <- RunUMAP(RasV_merged, reduction = "harmony", dims = 1:50, assay = "SCT", seed.use= 9999)
#p1 <- DimPlot(RasV_merged, reduction = "umap", group.by = "condition", shuffle = TRUE)
p2 <- DimPlot(RasV_merged, reduction = "umap", group.by = "batch", shuffle = TRUE)
p2
```
## Change cluster resolutions if i want

```{r}
RasV_merged <- FindNeighbors(RasV_merged, reduction = "harmony", dims = 1:50) 
RasV_merged <- FindClusters(RasV_merged, algorithm = 4, method = "igraph", resolution = c(0.06, 0.1, 0.08) )
RasV_merged <- RunUMAP(RasV_merged, reduction = "harmony", dims = 1:50, assay = "SCT", seed.use= 9999)


p4 <- DimPlot(RasV_merged, reduction = "umap", group.by = "SCT_snn_res.0.06", label = TRUE)
p4
ggsave("SCT/Individ_procc_Harmony.pdf", p4)
```

```{r}
######## Small density plot for exploring counts (here is per condition) #########

conditions <- unique(RasV_merged@meta.data$condition)


df_sct_counts <- data.frame()


for (cond in conditions) {

  cells_in_condition <- WhichCells(RasV_merged, expression = condition == cond)
  sct_data <- GetAssayData(RasV_merged, assay = "SCT", slot = "data")[, cells_in_condition]
  total_sct_counts <- Matrix::colSums(sct_data)
  

  temp_df <- data.frame(counts = total_sct_counts, Assay = "SCT", Condition = cond)
  df_sct_counts <- rbind(df_sct_counts, temp_df)
}


ggplot(df_sct_counts, aes(x = counts, fill = Condition)) +
  geom_density(alpha = 0.5) +
  labs(title = "Distribution of Total Counts per Cell by Condition",
       x = "Total Counts per Cell",
       y = "Density") +
  scale_fill_brewer(palette = "Set1") +
  theme_minimal()
```

## plot to get an idea of the cluster and the counts
```{r}

p5 <- FeaturePlot(object = RasV_merged, features = "nCount_RNA")
p6 <- FeaturePlot(object = RasV_merged, features = "nFeature_RNA")

p5/p6
```

## Clean the columns from metadata that we dont need anymore

```{r}

# remove the columns i dont want from the final df
columns_to_remove <- grepl("^DF\\.|^pANN", names(RasV_merged@meta.data))
RasV_merged@meta.data <- RasV_merged@meta.data[, !columns_to_remove]

#### We can join the RNA layers or keep them seperate##########
#RasV_merged[["RNA"]] <- JoinLayers(RasV_merged[["RNA"]], layers = "^data",new = "data")

head(RasV_merged@meta.data)
```


## maybe split the assay data maybe not

```{r}
# # Setting assay data after merging
# RasV_merged <- RasV_merged %>%
#   SetAssayData(assay = "SCT", slot = "data", function(x) {
#   # Check if 'batch' column is available
#   if("batch" %in% colnames(x@meta.data)) {
#     split(RasV_merged[["SCT"]], f = x@meta.data$batch)
#   } else {
#     stop("Batch column not found in meta.data after merging")
#   }
# })
```

## Change cluster resolutions if i want and save final version of integration

```{r}
# save final version
saveRDS(RasV_merged,"SCT/harmony_integrated_individ.rds" )
#RasV_merged <- readRDS("SCT/integrated_merged.rds" )

```

## seurat_cluster changes everytime i run the fund clusters but the previous resolutions are save in other columns





# Use ScVI  : https://docs.scvi-tools.org/en/1.0.0/tutorials/notebooks/scvi_in_R.html

```{r}



```

# Use seurats anchor method

