---
title: "Sc_Analysis_after_Qc_and_doublet"
author: "theodoros foskolos"
date: "2024-03-27"
output: html_document
---

## Libraries

```{r}
knitr::opts_chunk$set(echo = TRUE, fig.show = 'hide', results = 'hide')
```

library(Seurat)

```{r}
library(tidyverse)
library(Seurat)
library(patchwork)
library(presto)
library(DoubletFinder)
library(sctransform)
library(harmony)
library(dittoSeq)
#devtools::install_github("GreenleafLab/ArchR", ref="master", repos = BiocManager::repositories())
library(ArchR)
library(ggpubr)
library(magrittr)
library(RColorBrewer)
#ArchR::installExtraPackages()
set.seed(9999)

#library(Azimuth)

```

### mouse atlas: <https://github.com/ggjlab/scMCA> or Manual annotation or anchors or R

## input the merged integrated processed dataset

```{r}
#Sc_filtered<- readRDS("\SCT\intgr_indiv_SCT\data\")
Sc_filtered <- readRDS("SCT/intgr_indiv_SCT_harmony/data/annotated_intgr.rds")
#Sc_filtered$seurat_clusters <- RasV_merged$SCT_snn_res.0.06 #add here the one that you want as main cluster 
#DefaultAssay(object = Sc_filtered) <- "RNA"
```

## Using the SCT assay- DEG TOP CELL MARKERS

```{r}
data(Sc_filtered)
Sc_filtered <- Seurat::UpdateSeuratObject(Sc_filtered)

# here we find markers via presto by DEG of 1 cluster vs all the others. Important to use the resolution (clustering we want)
### !!!!!!! seurat clusters always get the last resolution from findClusters!!! so use that or the specific resolution you want to find markers!!!!!!###############

dif_analysis <- wilcoxauc(Sc_filtered, seurat_assay = "SCT","SCT_snn_res.0.06")

```

### Seems i need to join the RNA layers( i have one layer per experiment/ repli), ALSO SCT seems to remove some genes that may exist in less than 5 cells (vst = 5 default)

## Upregulated markers

```{r}
# set filtering params
top_genes <- top_markers(dif_analysis, n = 50, auc_min = .5, padj_max = 0.05, pct_in_min = 70)

top_genes <- top_genes[, sort(names(top_genes))]
write.csv(top_genes, "SCT/intgr_indiv_SCT/Cell_Markers_res_0_6.csv", row.names = FALSE)
```

## Extra resolutions if needed

```{r}
# Sc_filtered <- FindClusters(Sc_filtered, algorithm = 4, method = "igraph", resolution = 0.3 )
# Sc_filtered <- RunUMAP(Sc_filtered, reduction = "harmony", dims = 1:50, assay = "SCT", seed.use= 9999)
# 
p<- DimPlot(Sc_filtered, reduction = "umap", label = T, group.by = "cell_type", alpha = 0.14)+ ggtitle("Manual cell annotation - res 0.06")
p
ggsave("manual_celltype_annotation_res0_06.pdf")
# p
# ggsave("conditions_harmony_indivproc_res0_12c.pdf", p)
saveRDS(Sc_filtered,"SCT/harmony_integrated_individ.rds")
```

```{r}
Sc_subset <- subset(Sc_filtered, subset = batch %in% c("end_1", "end_2", "end_3"))

# Create a UMAP plot for the subsetted data
p4 <- DimPlot(Sc_subset, reduction = "umap", shuffle = TRUE, group.by = "batch", alpha = 0.1)
p4

Sc_subset <- SCTransform(Sc_subset, vars.to.regress = "percent.mt")
Sc_subset <- RunPCA(SCT_sub, assay = "SCT")
Sc_subset <- RunUMAP(Sc_subset, dims = 1:50, assay = "SCT")
DimPlot(Sc_subset, reduction = "umap", shuffle = TRUE, group.by = "batch", alpha = 0.1)

```

## Manual annotation of the cell types based on diff expression done for markers

```{r}
# manual identification through the DEG
# last annotation is for res: 0.08 - 0.1 seems better
cluster_to_celltype <- c( "1" = "EC",
                          "2" = "T",
                          "3" = "B",
                          "4" = "M/MC",
                          "5" = "NK/T",
                          "6" = "HEP1/C",
                          "7" = "plasm/?",
                          "8" = "Gran",
                          "9" = "pDCs",
                          "10" = "HEP2/CC"
                          )

```

# Pass the cell names of the manual identification to the cell_type column depending the clustering resolution we want

```{r}
# Sc_filtered@meta.data$cell_type <- cluster_to_celltype[as.character(Sc_filtered@meta.data$SCT_snn_res.0.06)]
# 
# p3<- DimPlot(Sc_filtered, reduction = "umap", label = T, group.by = "SCT_snn_res.0.06", alpha = 0.1)

cell_types <- levels(factor(Sc_filtered$cell_type))
col.ls <- ArchRPalettes[[4]][1:length(cell_types)]
names(col.ls) <- cell_types

# p4 <- DimPlot(Sc_filtered, reduction = "umap", shuffle = T, group.by = "cell_type", alpha= 0.1, label = T, cols = col.ls )
# p4
# saveRDS(Sc_filtered,"SCT/intgr_indiv_SCT_harmony/data/annotated_intgr.rds" )
```

## PLOTTING DISTRIVUTIONS

```{r}

stacked <- dittoBarPlot(
            object = Sc_filtered,
            var = "cell_type",
            group.by = "condition",
            x.reorder= c(1,3,2),
            #var.labels.reorder = c(1,12,9,8,11,10,5,4,2,3,6,7),
            main = "                                  Cell Types Per Condition",
            legend.title = "Cell Type"
            )

stacked

ggsave("SCT/cell_type_conditions_stackedbar.pdf", stacked, width = 10, height = 8)
```

```{r}


metadata <- Sc_filtered@meta.data


# look Better how to do these calculations on paper: also see the ggbarplot
# percentages per dataset of each celltype

percentage_data <- metadata %>%
  group_by(batch, cell_type, condition) %>%
  summarise(total_cells = n(), .groups = 'drop') %>%
  group_by(batch) %>%
  mutate(percentage = total_cells / sum(total_cells) * 100) %>%
  ungroup()

# calculate SD and mean

cell_perc <- percentage_data %>%
  group_by(condition, cell_type) %>%
  summarise(mean_percentage = mean(percentage),
            sd_percentage = sd(percentage),
            .groups = 'drop') %>%
  mutate(cell_type = as.factor(cell_type))

# order them descending
cell_perc <- cell_perc %>%
  group_by(cell_type) %>%
  mutate(avg_percentage = mean(mean_percentage)) %>%
  ungroup() %>%
  mutate(cell_type = reorder(cell_type, -avg_percentage)) # - for descending 


cell_types <- sort(unique(cell_perc$cell_type))
col.ls <- ArchRPalettes[1:length(cell_types)]
names(col.ls) <- cell_types


cell_types <- levels(factor(cell_perc$cell_type))

# error bars

error_bars <- cell_perc %>%
  arrange(condition, desc(cell_type)) %>%
  group_by(condition) %>%
  mutate(cumulative_mean = cumsum(mean_percentage)) %>%
  ungroup()

cell_perc <- cell_perc %>%
  mutate(condition = factor(condition, levels = c("ctl", "int", "end")))
```

## Stacked bar plot

```{r}

col.ls <- ArchRPalettes[[4]][1:length(cell_types)] ## change the first number in archR to change palletes
names(col.ls) <- cell_types


p <- ggplot(cell_perc, aes(x = condition, y = mean_percentage, fill = cell_type)) +
  geom_bar(stat = 'identity', position = 'stack') +
  geom_errorbar(
    data = error_bars,
    aes(x = condition, ymin = cumulative_mean - sd_percentage, ymax = cumulative_mean + sd_percentage, group = cell_type),
    width = 0.3,
    position = position_dodge(0.2)
  ) +
  scale_fill_manual(values = col.ls) +
  labs(y = "Average Percentage", x = "Condition", fill = "Cell Type") +
  theme_minimal()

ggsave("SCT/cell_type_manual_0_08.pdf", p, width = 10, height = 8)
p
```

## Separate bar plots per cell type

```{r}


# plot all cell types per condition
p1 <- ggplot(cell_perc, aes(x = condition, y = mean_percentage, fill = condition)) +
      geom_bar(stat = 'identity', position = position_dodge(), width = 0.7) +
      geom_errorbar(aes(ymin = mean_percentage - sd_percentage, ymax = mean_percentage + sd_percentage),
      width = 0.2, position = position_dodge(0.7)) +
      facet_wrap(~cell_type) +
      labs(x = "Condition", y = "Percentage", title = "Cell Type per stage") +
      theme_dark()
p1
ggsave("output/cell_type_0_06.pdf", p1, width = 10, height = 8)
```

## B) Label Transfer - from human annotated file

## I make some subsets for time (remove it after) I need to first preprocess my refference data-set. Need to preprocess it in the same way i did my query data-set.

```{r}
# load the refference dataset
load("data/annotated_file/Xue.subsampled_sco.RData")


#Sc_filtered  <-readRDS("SCT/annotated_intgr_res_0_1.rds")
```

```{r}
subset_sample <- rownames(Xue.subsampled_sco@meta.data)[Xue.subsampled_sco@meta.data$Cancer_type == "HCC"]

ref_annot <- subset(Xue.subsampled_sco, cells = subset_sample)
```

```{r}
nCount_values <- ref_annot@meta.data$nCount_RNA

#  upper and lower quantile limits
upper_lim_nCount <- quantile(nCount_values, 0.99)
lower_lim_nCount <- quantile(nCount_values, 0.01)

ref_annot <- subset(ref_annot, subset = nFeature_RNA > 400 & 
                                            percent.mt < 10 & 
                                            nCount_RNA > lower_lim_nCount & 
                                            nCount_RNA < upper_lim_nCount)
```

## Get less cells if we want faster calculations ( not recommended )

```{r}
num_cells_to_sample <- ceiling(length(ref_annot@meta.data$Cancer_type) / 2)

random_cells <- sample(rownames(ref_annot@meta.data), num_cells_to_sample)

# 3. Subset ref_annot with the selected cells
ref_annot <- subset(ref_annot, cells = random_cells)
```

# Anchor the 2 datasets

```{r}
DefaultAssay(ref_annot) <- 'SCT'
DefaultAssay(Sc_filtered) <- 'SCT'

Sc_filtered <- PrepSCTFindMarkers(Sc_filtered)
anchors <- FindTransferAnchors(reference = ref_annot, query = Sc_filtered, dims = 1:50, reference.assay = "SCT", query.assay = "SCT", reduction = "pcaproject" )
```

```{r}
predicted.labels <- TransferData(
  anchorset = anchors,
  refdata = ref_annot@meta.data$clusters,
  dims = 1:50
)
```

```{r}
target.data <- AddMetaData(object = Sc_filtered, metadata = predicted.labels$predicted.id, col.name = "predicted_label")
```

```{r}
#we can set a threshold here
hist(predicted.labels$prediction.score.max)
```

```{r}
Sc_filtered <- AddMetaData(object = Sc_filtered, metadata = predicted.labels$predicted.id, col.name = "anchor_pred")
```

```{r}
# if i want to makje broader groups ( eg. CD4_dkjbvjsbv -> CD4)
Sc_filtered$celltype_wide <- gsub("_.*", "", Sc_filtered$anchor_pred)
```

```{r}

umapp <- DimPlot(Sc_filtered, reduction = "umap", group.by = "celltype_wide", label = T, alpha = 0.1)
umapp

ggsave("SCT/Xue_annot.pdf", dpi= 300)
```

## if i want to find more DEG from top markers like negative etc.

```{r}
# Add the LOGFC value next to the top_markers
append_logFC_by_celltype <- function(cell_type_column, cell_type_name, diff_exp_data) {
  # Filter the differential expression data for the current cell type
  diff_exp_subset <- diff_exp_data[diff_exp_data$group == cell_type_name, ]
  
  sapply(cell_type_column, function(gene_name) {
    matched_logFC <- diff_exp_subset[diff_exp_subset$feature == gene_name, "logFC"]
    if (length(matched_logFC) != 0) {
      return(paste0(gene_name, " (", matched_logFC, ")"))
    } else {
      return(gene_name)
    }
  })
}


final_topgenes <- as.data.frame(mapply(append_logFC_by_celltype, top_genes, colnames(top_genes), MoreArgs = list(diff_exp_data = dif_analysis)))

```

```{r}
# Top 40 positive
top_positive <- dif_analysis %>%
  group_by(group) %>%
  filter(padj > 0.05) %>%
  slice_max(order_by = logFC, n = 50)

# Top 40 negative
top_negative <- dif_analysis %>%
  group_by(group) %>%
  filter(padj > 0.05) %>% # put more filters if needed
  slice_min(order_by = logFC, n = 50)

```

```{r}
pos <- top_positive %>%
  select(group, feature) %>% # Select only the necessary columns
  #pivot_wider(names_from = group, values_from = feature)

```

```{r}
top_genes <- dif_analysis_filt %>%
  group_by(celltype) %>%
  arrange(desc(logFC)) %>%
  top_n(20, wt = logFC) %>%
  ungroup() %>%
  bind_rows(
    dif_analysis_filt %>%
    group_by(celltype) %>%
    arrange(logFC) %>%
    top_n(20, wt = -logFC) %>%
    ungroup()
  )

print(top_genes)
```

```{r}
features_to_plot <- intersect(rownames(Sc_filtered), top_genes$feature)

# Visualize top genes using FeaturePlot or DoHeatmap
# For FeaturePlot
FeaturePlot(Sc_filtered, features = features_to_plot)

# For DoHeatmap
DoHeatmap(Sc_filtered, features = features_to_plot)
```

```{r}
df_wide <- dif_analysis_filt %>%
  pivot_wider(
    names_from = celltype,  # this is your cell type column
    values_from = feature,  # this is your differentially expressed genes column
    values_fill = list(feature = NA)  # How to fill in missing cases
  )
df_wide
```
