---
title: "CC_inter"
output: html_document
date: "2024-04-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

 
BiocManager::install("ComplexHeatmap")
BiocManager::install("limma")
devtools::install_github("saeyslab/nichenetr")
```

```{r}
devtools::install_github("saeyslab/nichenetr")
```

```{r}
library(nichenetr)
library(Seurat)
library(SeuratObject)
library(tidyverse)
```

## Niche net - ligand receptor predictions- look here :

## <https://github.com/saeyslab/nichenetr/blob/master/vignettes/seurat_steps.md#receptors-of-top-ranked-ligands>

and for use cases:

## <https://github.com/saeyslab/nichenetr/blob/master/vignettes/faq.md>

```{r}
Sc_filtered <- readRDS(file ="SCT/intgr_indiv_SCT_harmony/Subclustering/subcluster_annotated_RasV.rds")
DefaultAssay(Sc_filtered) <-"SCT"
```

```{r}
Sc_filtered@meta.data %>% head()
```

```{r}
Sc_filtered <- UpdateSeuratObject(Sc_filtered)
```

```{r}
DimPlot(Sc_filtered, reduction = "umap", group.by = "cell_subtype")
```

```{r}
organism = "mouse"

if(organism == "human"){
  lr_network = readRDS(url("https://zenodo.org/record/7074291/files/lr_network_human_21122021.rds"))
  ligand_target_matrix = readRDS(url("https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final.rds"))
  weighted_networks = readRDS(url("https://zenodo.org/record/7074291/files/weighted_networks_nsga2r_final.rds"))
} else if(organism == "mouse"){
  lr_network = readRDS(url("https://zenodo.org/record/7074291/files/lr_network_mouse_21122021.rds"))
  ligand_target_matrix = readRDS(url("https://zenodo.org/record/7074291/files/ligand_target_matrix_nsga2r_final_mouse.rds"))
  weighted_networks = readRDS(url("https://zenodo.org/record/7074291/files/weighted_networks_nsga2r_final_mouse.rds"))

}

lr_network = lr_network %>% distinct(from, to)
head(lr_network)
```

```{r}
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network, by = c("from","to"))
head(weighted_networks$lr_sig) # interactions and their weights in the ligand-receptor + signaling network
```

```{r}
head(weighted_networks$gr) # interactions and their weights in the gene regulatory network
```

```{r}
#if the dataset has older gene symbols we need to do something like this 
#Sc_filtered = alias_to_symbol_seurat(Sc_filtered, "mouse")
```

```{r}
# in my annotated file  i have all the subtypes so i can make a generalized celltype column 

Sc_filtered$cell_type_gen <- gsub("_.*", "", Sc_filtered$cell_subtype)
Sc_filtered$cell_type_gen <- gsub("^[^_]*_", "", Sc_filtered$cell_subtype)
Sc_filtered$cell_type_gen <- Sc_filtered$cell_subtype
```

```{r}
# Set the identity classes to your celltypes column
Idents(Sc_filtered) <- Sc_filtered$cell_type_gen
```

## 1. Define a “sender/niche” cell population and a “receiver/target” cell population present in your expression data and determine which genes are expressed in both populations

```{r}

## receiver
receiver = "Dysf/Exhausted_CD8+_T"
expressed_genes_receiver = get_expressed_genes(receiver, Sc_filtered, pct = 0.10)# expressed in 10% of the cells

background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]

```

```{r}
# unique(Sc_filtered$cell_type_gen) - look at the general cell clusters
## sender, add the sender types
sender_celltypes = c("Cxcl16+Apoe+_Mph")

list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, Sc_filtered, 0.10) # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
```

## 2. Define a gene set of interest: these are the genes in the “receiver/target” cell population that are potentially affected by ligands expressed by interacting cells (e.g. genes differentially expressed upon cell-cell interaction)

```{r}
seurat_obj_receiver= subset(Sc_filtered, idents = receiver)
seurat_obj_receiver = SetIdent(seurat_obj_receiver, value = seurat_obj_receiver[["condition", drop=TRUE]])

#add the conditions here
condition_oi = "int"
condition_reference = "ctl"  ## maybe i can add sthe 3rd condition 2

PrepSCTFindMarkers(Sc_filtered, assay = "SCT", verbose = TRUE) 
DE_table_receiver = FindMarkers(object = seurat_obj_receiver, ident.1 = condition_oi, ident.2 = condition_reference, min.pct = 0.10, recorrect_umi=FALSE) %>% rownames_to_column("gene")## recorrect umi needed!!

geneset_oi = DE_table_receiver %>% filter(p_val_adj <= 0.05 & abs(avg_log2FC) >= 0.25) %>% pull(gene)
geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```

```{r}
# seurat_obj_receiver[["RNA"]] <- as(object = seurat_obj_receiver[["RNA"]], Class = "Assay")
# pseudo_bulk <- AggregateExpression(seurat_obj_receiver, assays = "RNA", return.seurat = TRUE, group.by = c("int","ctl","cell_subtype"))
# # each 'cell' is a donor-condition-celltype pseudobulk profile
# tail(Cells(pseudo_bulk))
# 
# Idents(pseudo_bulk) <- "cell_subtype"
# 
# bulk.mono.de <- FindMarkers(object = pseudo_counts, 
#                          test.use = "DESeq2",ident.1 = "Naive_T")
# head(bulk.mono.de, n = 15)


```

## 3. Define a set of potential ligands: these are ligands that are expressed by the “sender/niche” cell population and bind a (putative) receptor expressed by the “receiver/target” population

```{r}
ligands = lr_network %>% pull(from) %>% unique()
receptors = lr_network %>% pull(to) %>% unique()

expressed_ligands = intersect(ligands,expressed_genes_sender)
expressed_receptors = intersect(receptors,expressed_genes_receiver)

potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()

```

## 4) Perform NicheNet ligand activity analysis: rank the potential ligands based on the presence of their target genes in the gene set of interest (compared to the background set of genes)

```{r}

ligand_activities = predict_ligand_activities(geneset = geneset_oi, background_expressed_genes = background_expressed_genes, ligand_target_matrix = ligand_target_matrix, potential_ligands = potential_ligands)

ligand_activities = ligand_activities %>% arrange(-aupr_corrected) %>% mutate(rank = rank(dplyr::desc(aupr_corrected)))
ligand_activities
```

```{r}
# top-ranked ligands
best_upstream_ligands = ligand_activities %>% top_n(30, aupr_corrected) %>% arrange(-aupr_corrected) %>% pull(test_ligand) %>% unique()


```

```{r}
## plot the top 30 genes for expressed ligands
DotPlot(Sc_filtered, features = best_upstream_ligands %>% rev(), cols = "RdYlBu") + RotatedAxis()

```

## 5) Infer receptors and top-predicted target genes of ligands that are top-ranked in the ligand activity analysis

```{r}
# Active target gene inference

active_ligand_target_links_df = best_upstream_ligands %>% lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>% bind_rows() %>% drop_na()

active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.33)

order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
order_targets = active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()
rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23

vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()
```

```{r}
p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple",legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential")  + theme(axis.text.x = element_text(face = "italic")) + scale_fill_gradient2(low = "whitesmoke",  high = "purple", breaks = c(0,0.0045,0.0090))
p_ligand_target_network
## Note that not all ligands from the top 30 are present in this ligand-target heatmap. The left-out ligands are ligands that don’t have target genes with high enough regulatory potential scores. Therefore, they did not survive the used cutoffs. To include them, you can be less stringent in the used cutoffs.
```

### Receptors of top-ranked ligands

```{r}
lr_network_top = lr_network %>% filter(from %in% best_upstream_ligands & to %in% expressed_receptors) %>% distinct(from,to)
best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()

lr_network_top_df_large = weighted_networks_lr %>% filter(from %in% best_upstream_ligands & to %in% best_upstream_receptors)

lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
lr_network_top_matrix = lr_network_top_df %>% select(-to) %>% as.matrix() %>% magrittr::set_rownames(lr_network_top_df$to)

dist_receptors = dist(lr_network_top_matrix, method = "binary")
hclust_receptors = hclust(dist_receptors, method = "ward.D2")
order_receptors = hclust_receptors$labels[hclust_receptors$order]
    
dist_ligands = dist(lr_network_top_matrix %>% t(), method = "binary")
hclust_ligands = hclust(dist_ligands, method = "ward.D2")
order_ligands_receptor = hclust_ligands$labels[hclust_ligands$order]

order_receptors = order_receptors %>% intersect(rownames(lr_network_top_matrix))
order_ligands_receptor = order_ligands_receptor %>% intersect(colnames(lr_network_top_matrix))

vis_ligand_receptor_network = lr_network_top_matrix[order_receptors, order_ligands_receptor]
rownames(vis_ligand_receptor_network) = order_receptors %>% make.names()
colnames(vis_ligand_receptor_network) = order_ligands_receptor %>% make.names()
```

```{r}
p_ligand_receptor_network = vis_ligand_receptor_network %>% t() %>% make_heatmap_ggplot("Ligands","Receptors", color = "mediumvioletred", x_axis_position = "top",legend_title = "Prior interaction potential")
p_ligand_receptor_network
```

## 6) Add log fold change information of ligands from sender cells

```{r}
# DE analysis for each sender cell type
# this uses a new nichenetr function - reinstall nichenetr if necessary!

DE_table_all = Idents(Sc_filtered) %>% levels() %>% intersect(sender_celltypes) %>% lapply(PrepSCTFindMarkers,Sc_filtered) %>% lapply(get_lfc_celltype, seurat_obj = Sc_filtered, condition_colname = "condition", condition_oi = condition_oi, condition_reference = condition_reference, expression_pct = 0.10, celltype_col = "cell_type_gen") %>% reduce(full_join) # use this if cell type labels are the identities of your Seurat object -- if not: indicate the celltype_col properly,recorrect_umi=FALSE
DE_table_all[is.na(DE_table_all)] = 0

# Combine ligand activities with DE information
ligand_activities_de = ligand_activities %>% select(test_ligand, pearson) %>% rename(ligand = test_ligand) %>% left_join(DE_table_all %>% rename(ligand = gene))
ligand_activities_de[is.na(ligand_activities_de)] = 0

# make LFC heatmap
lfc_matrix = ligand_activities_de  %>% select(-ligand, -pearson) %>% as.matrix() %>% magrittr::set_rownames(ligand_activities_de$ligand)
rownames(lfc_matrix) = rownames(lfc_matrix) %>% make.names()

order_ligands = order_ligands[order_ligands %in% rownames(lfc_matrix)]
vis_ligand_lfc = lfc_matrix[order_ligands,]

colnames(vis_ligand_lfc) = vis_ligand_lfc %>% colnames() %>% make.names()

p_ligand_lfc = vis_ligand_lfc %>% make_threecolor_heatmap_ggplot("Prioritized ligands","LFC in Sender", low_color = "midnightblue",mid_color = "white", mid = median(vis_ligand_lfc), high_color = "red",legend_position = "top", x_axis_position = "top", legend_title = "LFC") + theme(axis.text.y = element_text(face = "italic"))
p_ligand_lfc
```

```{r}
# change colors a bit to make them more stand out
p_ligand_lfc = p_ligand_lfc + scale_fill_gradientn(colors = c("midnightblue","blue", "grey95", "grey99","firebrick1","red"),values = c(0,0.1,0.2,0.25, 0.40, 0.7,1), limits = c(vis_ligand_lfc %>% min() - 0.1, vis_ligand_lfc %>% max() + 0.1))
p_ligand_lfc
```

## 7) Summary visualizations of the NicheNet analysis

```{r}
# ligand activity heatmap
ligand_aupr_matrix = ligand_activities %>% select(aupr_corrected) %>% as.matrix() %>% magrittr::set_rownames(ligand_activities$test_ligand)

rownames(ligand_aupr_matrix) = rownames(ligand_aupr_matrix) %>% make.names()
colnames(ligand_aupr_matrix) = colnames(ligand_aupr_matrix) %>% make.names()

vis_ligand_aupr = ligand_aupr_matrix[order_ligands, ] %>% as.matrix(ncol = 1) %>% magrittr::set_colnames("AUPR")
p_ligand_aupr = vis_ligand_aupr %>% make_heatmap_ggplot("Prioritized ligands","Ligand activity", color = "darkorange",legend_position = "top", x_axis_position = "top", legend_title = "AUPR\n(target gene prediction ability)") + theme(legend.text = element_text(size = 9))
```

```{r}
# ligand expression Seurat dotplot
order_ligands_adapted <- str_replace_all(order_ligands, "\\.", "-")
rotated_dotplot = DotPlot(Sc_filtered %>% subset(celltype %in% sender_celltypes), features = order_ligands_adapted, cols = "RdYlBu") + coord_flip() + theme(legend.text = element_text(size = 10), legend.title = element_text(size = 12)) # flip of coordinates necessary because we want to show ligands in the rows when combining all plots
```

```{r}
figures_without_legend = cowplot::plot_grid(
  p_ligand_aupr + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()),
  rotated_dotplot + theme(legend.position = "none", axis.ticks = element_blank(), axis.title.x = element_text(size = 12), axis.text.y = element_text(face = "italic", size = 9), axis.text.x = element_text(size = 9,  angle = 90,hjust = 0)) + ylab("Expression in Sender") + xlab("") + scale_y_discrete(position = "right"),
  p_ligand_lfc + theme(legend.position = "none", axis.ticks = element_blank()) + theme(axis.title.x = element_text()) + ylab(""),
  p_ligand_target_network + theme(legend.position = "none", axis.ticks = element_blank()) + ylab(""),
  align = "hv",
  nrow = 1,
  rel_widths = c(ncol(vis_ligand_aupr)+6, ncol(vis_ligand_lfc) + 7, ncol(vis_ligand_lfc) + 8, ncol(vis_ligand_target)))

legends = cowplot::plot_grid(
    ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_aupr)),
    ggpubr::as_ggplot(ggpubr::get_legend(rotated_dotplot)),
    ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_lfc)),
    ggpubr::as_ggplot(ggpubr::get_legend(p_ligand_target_network)),
    nrow = 1,
    align = "h", rel_widths = c(1.5, 1, 1, 1))

combined_plot = cowplot::plot_grid(figures_without_legend, legends, rel_heights = c(10,5), nrow = 2, align = "hv")
combined_plot
```
